### 2019-05-19 source/dialogues.tex

\begin{table}[H]
\centering
\resizebox{\textwidth}{!}{
\begin{tabular}[]{lll}
\toprule

turn&$\beta=0$&$\beta=0.5$\tabularnewline
\midrule
turn 0& \makecell[l]{valid slots : [0, 0, 0]\\ srs : [ None None None ]\\ system says ASK\_ORAL(1) \\ user says INFORM} &\makecell[l]{ valid slots : [0, 0, 0]\\ srs : [ None None None ]\\ system says ASK\_ORAL(2)\\ user says INFORM}\tabularnewline\midrule
turn 1&\makecell[l]{valid slots : [0, 0, 0]\\ srs : [ None 0.48 None ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1]\\ srs : [ None None 0.56 ]\\ system says ASK\_ORAL(0)\\ user says INFORM}\tabularnewline\midrule
turn 2&\makecell[l]{valid slots : [0, 0, 0]\\ srs : [ None 0.48 0.22 ]\\ system says ASK\_ORAL(0) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1] \\ srs : [ 0.30 None 0.56 ]\\ system says ASK\_ORAL(1)\\ user says INFORM}\tabularnewline\midrule
turn 3&\makecell[l]{valid slots : [0, 0, 0]\\ srs : [ 0.62 0.48 0.22 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1]\\ srs : [ 0.30 0.54 0.56 ]\\ system says ASK\_ORAL(0)\\ user says INFORM}\tabularnewline\midrule
turn 4&\makecell[l]{valid slots : [0, 0, 0] \\ srs : [ 0.62 0.48 0.66 ]\\ system says ASK\_ORAL(1) \\ user says INFORM} & \makecell[l]{valid slots : [0, 0, 1]\\ srs : [ 0.68 0.54 0.56 ]\\ system says ASK\_NUM\_PAD(1)\\ user says INFORM}\tabularnewline\midrule
turn 5&\makecell[l]{valid slots : [0, 1, 0]\\ srs : [ 0.62 0.56 0.66 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 0.56 ]\\ system says SUMMARIZE\_AND\_INFORM\\ user says DENY\_SUMMARIZE}\tabularnewline\midrule
turn 6&\makecell[l]{valid slots : [0, 1, 0]\\ srs : [ 0.62 0.56 0.14 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 0.56 ]\\ system says ASK\_NUM\_PAD(2)\\ user says INFORM}\tabularnewline\midrule
turn 7&\makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.62 0.56 0.30 ]\\ system says ASK\_ORAL(2) \\ user says INFORM }& \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 1.00 ]\\ system says SUMMARIZE\_AND\_INFORM\\ user says DENY\_SUMMARIZE}\tabularnewline\midrule
turn 8&\makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.62 0.56 0.49 ]\\ system says ASK\_ORAL(2) \\ user says INFORM} & \makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.68 1.00 1.00 ]\\ system says ASK\_NUM\_PAD(0)\\ user hangs up !}\tabularnewline\midrule
turn 9&\makecell[l]{valid slots : [0, 1, 1]\\ srs : [ 0.62 0.56 0.65 ]\\ system says SUMMARIZE\_AND\_INFORM \\ max size reached !}& \makecell[l]{ }\tabularnewline\bottomrule
\end{tabular}
}
    \caption{Two dialogues generated by a safe policy ($\beta=0$) on the left and a risky one ($\beta=0.5$) on the right.}
    \label{table:dialogues}
\end{table}