### 2019-05-21 main.tex
Updates from Overleaf

\documentclass{article}

% if you need to pass options to natbib, use, e.g.:
%     \PassOptionsToPackage{numbers, compress}{natbib}
% before loading neurips_2019

% ready for submission
% \usepackage{neurips_2019}

% to compile a preprint version, e.g., for submission to arXiv, add add the
% [preprint] option:
%     \usepackage[preprint]{neurips_2019}

% to compile a camera-ready version, add the [final] option, e.g.:
\usepackage{neurips_2019}

% to avoid loading the natbib package, add option nonatbib:
%     \usepackage[nonatbib]{neurips_2019}

\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc}    % use 8-bit T1 fonts
\usepackage{url}            % simple URL typesetting
\usepackage{booktabs}       % professional-quality tables
\usepackage{amsfonts}       % blackboard math symbols
\usepackage{nicefrac}       % compact symbols for 1/2, etc.
\usepackage{microtype}      % microtypography

% Paper specific includes and commands
\usepackage{times}
\usepackage{soul}
\usepackage[small]{caption}
\usepackage{graphicx}
\usepackage{amsmath}
\urlstyle{same}
\usepackage{xargs}
\usepackage{amsthm}
\usepackage{mathtools}
\usepackage{dsfont}
\usepackage{import}
\usepackage{svg}
\usepackage{pdfpages}
\usepackage[norelsize, linesnumbered, ruled, lined, boxed, commentsnumbered]{algorithm2e}
\usepackage{float}
\usepackage{tabularx}
\usepackage{makecell}
\usepackage[lofdepth,lotdepth]{subfig}

% Custom math commands
\input{mathdef}

% Todo notes
\usepackage[colorinlistoftodos,prependcaption,textsize=tiny]{todonotes}
\usetikzlibrary{patterns}
\newcommand{\quotes}[1]{``#1''}
\newcommandx{\nc}[2][1=]{\todo[linecolor=blue,backgroundcolor=blue!25,bordercolor=blue,#1]{\textbf{Nicolas:} #2}}
\newcommandx{\op}[2][1=]{\todo[linecolor=green,backgroundcolor=green!25,bordercolor=green,#1]{\textbf{Olivier:} #2}}
\newcommandx{\el}[2][1=]{\todo[linecolor=red,backgroundcolor=red!25,bordercolor=red,#1]{\textbf{Edouard:} #2}}
\newcommandx{\td}[2][1=]{\todo[inline,size=\large,#1]{#2}}

% Resources path
\makeatletter
\providecommand*{\input@path}{}
\g@addto@macro\input@path{{./source/}{./source/ressources/}{./ressources/}}
\makeatother

%---------------------------------------------
% Cross Referencing of Supplementary Material %---------------------------------------------
% https://tex.stackexchange.com/questions/14364/cross-referencing-between-different-files
% https://tex.stackexchange.com/questions/244239/correct-ordering-of-xr-hyper-hyperref-and-cleveref
% \usepackage{xr}
% \usepackage{xr-hyper}
\usepackage{hyperref}
\usepackage[capitalise,noabbrev]{cleveref}
% \makeatletter
% \newcommand*{\addFileDependency}[1]{% argument=file name and extension
%   \typeout{(#1)}
%   \@addtofilelist{#1}
%   \IfFileExists{#1}{}{\typeout{No file #1.}}
% }
% \makeatother
% \newcommand*{\myexternaldocument}[1]{%
%     \externaldocument{#1}%
%     \addFileDependency{#1.tex}%
%     \addFileDependency{#1.aux}%
% }
% \myexternaldocument{supplementary}

\title{Budgeted Reinforcement Learning in Continuous State Space}

\author{
Nicolas Carrara$^{1,}$\thanks{equal contribution}\and
Edouard Leurent$^{1,3,}$\footnotemark[1]\and
Romain Laroche$^{4}$\and
Tanguy Urvoy$^{2}$\and
Odalric Maillard$^{1}$\and
Olivier Pietquin$^{1,2,}$\thanks{now with Google Research - Brain Team, Paris, France.}\\
\affiliations $^1$SequeL team, INRIA Lille -- Nord Europe, France\\
$^2$Orange Labs, Lannion, France\\
$^3$Renault Group, France\\
$^4$Microsoft Research , Montreal, Canada
\emails
\{nicolas.carrara, edouard.leurent, odalric.maillard, olivier.pietquin\}@inria.fr,
romain.laroche@microsoft.com, tanguy.urvoy@orange.com
}

\begin{document}
\maketitle
\begin{abstract}
    A Budgeted Markov Decision Process (BMDP) is an extension of a Markov Decision Process to critical applications requiring safety constraints. It relies on a notion of risk implemented in the shape of a cost signal constrained  to lie below a threshold that -- importantly -- can be modified in real-time. So far, BMDPs could only be solved in the case of finite state spaces with known dynamics. This work extends the state-of-the-art to continuous spaces environments and unknown dynamics. We show that the solution to a BMDP is a fixed point of a novel Budgeted Bellman Optimality operator. This observation allows us to introduce natural extensions of Deep Reinforcement Learning algorithms to address large-scale BMDPs. We validate our approach on two simulated applications: spoken dialogue and autonomous driving.
\end{abstract}

% @Nicolas/Edouard: faites ici la liste des sections qui sont pretes a etre relues.
% Section 1: il manque les contributions
% Section 2: OK. On pense éventuellement rajouter une hypothèse de contractance de Tau (pas sûr que ce soit le cas)
% Section 3: OK
% Section 4: OK. Il faut aussi reprendre les notations des graphiques.
% Section 5: OK
% Section 6: OK
% Section 7: OK

\section{Introduction}

Reinforcement Learning (RL) is a general framework for decision-making under uncertainty. It frames the learning objective as the optimal control of a Markov Decision Process  $(\cS, \cA, P, R_r, \gamma)$ with continuous state space $\cS$, discrete actions $\cA$, and unknown dynamics $P\in \cM(\cS)^{\cS \times \cA}$ and rewards $R_r\in\Real^{\cS \times \cA}$, where $\mathcal{M}(\mathcal{X})$ denotes the set of probability measures over a measurable space $\mathcal{X}$. Formally, we seek a policy $\pi\in\cM(A)^\cS$ that maximises in expectation the $\gamma$-discounted return of rewards $G_r^\pi = \sum_{t=0}^\infty \gamma^t R_r(s_t, a_t)$.

However, this modelling assumption comes at a price: no control is given over the spread of the performance distribution. In many critical real-world applications where failures may turn out very costly, this is an issue as most decision-makers would rather give away some amount of expected optimality to increase the performances in the lower-tail of the distribution. This has led to the development of several risk-averse variants where the optimisation criteria include other statistics of the performance, such as the worst-case realisation \citep{Iyengar2005,Nilim2005,Wiesemann2013}, the variance-penalised expectation \citep{Garcia2015,Tamar2012}, or the Conditional Value-At-Risk (CVaR) \citep{Chow2014,DBLP:journals/corr/ChowGJP15}.

Reinforcement Learning also assumes that the performance can be described solely by a reward function $R_r$. Conversely, real problems typically involve many aspects, some of which can be contradictory \citep{ChunmingLiu2014}. For instance, a self-driving car needs to balance between progressing quickly on the road and avoiding collisions. When aggregating several objectives in a single scalar signal, as often in Multi-Objectives RL~\citep{Roijers2013ASO}, no control is given over their relative ratios, as high rewards can compensate high penalties. For instance, if a weighted sum is used to balance velocity $v$ and accidents $a$, then for any given choice of weights $\omega$ the optimality equation $\omega_v\expectedvalue[\sum\gamma^t v_t] + \omega_a\expectedvalue[\sum\gamma^t a_t] = G_r^* = \max_\pi G^\pi_r$ is the equation of a line in $(\expectedvalue[\sum\gamma^t v_t], \expectedvalue[\sum\gamma^t a_t])$, and the automotive company cannot control where its optimal policy $\pi^*$ lies on that line.

Both of these concerns can be addressed in the \emph{Constrained Markov Decision Process} (CMDP) setting \citep{BEUTLER1985236,Altman95constrainedmarkov}. In this multi-objective formulation, task completion and safety are considered separately. We equip the MDP with a cost signal $R_c \in \Real^{\cS\times \cA}$ and a cost budget $\beta\in\Real$. Similarly to $G_r$, we define the return of costs $G_c^\pi = \sum_{t=0}^\infty \gamma^t R_c(s_t, a_t)$ and the new cost-constrained objective:
\begin{equation}
\label{eq:cmdp}
\begin{array}{lcr}
 \max_{\pi\in\cM(\cA)^\cS} \expectedvalue[G_r^\pi | s_0=s, \pi] & \text{ s.t. } & \expectedvalue[G_c^\pi | s_0=s, \pi] \leq \beta
\end{array}
\end{equation}

This constrained framework allows for better control of the performance-safety tradeoff. However, it suffers from a major limitation: the budget has to be chosen before training, and cannot be changed afterwards.

To address this concern, the \emph{Budgeted Markov Decision Process} (BMDP) was introduced in \citep{Boutilier_Lu:uai16} as an extension of CMDPs to enable the online control over the budget $\beta$ within an interval $\cB \subset \Real$ of admissible budgets. Instead of fixing the budget prior to training, the objective is now to find a generic optimal policy $\pi^*$ that takes $\beta$ as input so as to solve the corresponding CMDP \eqref{eq:cmdp} for all $\beta\in\cB$. This gives the system designer the ability to move the optimal policy $\pi^*$ in real-time along the Pareto-optimal curve of the different reward-cost trade-offs.

Our first contribution is to re-frame the original BMDP formulation in the context of continuous states and infinite discounted horizon. We then propose a novel Budgeted Bellman Optimality Operator and prove the optimal value function to be a fixed point of this operator. Second, we use this operator in a Reinforcement Learning algorithm for solving BMDPs online by interaction with an environment, through function approximation and a tailored exploration procedure. Third, we scale this algorithms to large problems by providing an efficient implementation of the Budgeted Bellman Optimality Operator based on convex computing, and by leveraging tools from Deep Reinforcement Learning such as Deep Neural Networks and synchronous parallel computing. Finally, we validate our approach in two environments that display a clear trade-off between rewards and costs: a spoken dialogue system and a problem of behaviour planning for autonomous driving.

The proofs of our main results are provided in Supplementary Material \ref{sec:proofs}.

\section{Budgeted Dynamic Programming}

We will work in the space of budgeted policies, where $\pi$ both depends on $\beta$ and also outputs a next budget $\beta_a$. Hence, the budget $\beta$ is neither fixed nor constant as in the CMDP setting but instead evolves as part of the dynamics.

For ease of notation, we consider the \emph{augmented} state and action spaces $\ocS = \cS\times \cB$ and $\ocA= \cA\times \cB$, and equip them with the augmented dynamics $\ov{P}\in \cM(\ocS)^{\ocS \times \ocA}$ defined as:
\begin{equation*}
    \ov{P}\left(\os' \condbar \os, \oa\right) = \ov{P}\left((s',\beta') \condbar (s,\beta), (a, \beta_a)\right) \eqdef P(s'|s, a)\delta(\beta' - \beta_a),
\end{equation*}
where $\delta$ is the Dirac distribution.

In other words, in these augmented dynamics, the output budget $\beta_a$ returned at time $t$ by a budgeted policy $\pi\in \Pi=\cM(\ocA)^{\ocS}$ will be used to condition the policy at the next timestep $t+1$.

We stack the rewards and cost functions in a single \emph{vectorial} signal $R \in (\Real^2)^{{\ocS \times \ocA}}$.
Given an augmented transition $(\os, \oa) =((s,\beta), (a, \beta_a))$, we define:
\begin{equation}
     R(\os, \oa) \eqdef  \begin{bmatrix}
     R_r(s, a)\\
     R_c(s, a)
     \end{bmatrix}\in\Real^2.
\end{equation}

As earlier, the return $G^\pi = (G_r^\pi, G_c^\pi)$ of a budgeted policy $\pi\in\Pi$ refers to:
    $G^\pi \eqdef \sum_{t=0}^\infty \gamma^t R(\os_t, \oa_t)$.

We can introduce the value function $V^\pi$ and action-value function $Q^\pi$ of a budgeted policy $\pi\in\Pi$ as:
\begin{equation}
    \label{eq:value-function}
V^\pi(\os) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os\right] \qquad Q^\pi(\os, \oa) \eqdef \expectedvalue\left[ G^\pi \condbar \ov{s_0} = \os, \ov{a_0} = \oa\right].
\end{equation}

Likewise, we note $V^\pi = (V_r^\pi, V_c^\pi)$ and $Q^\pi = (Q_r^\pi, Q_c^\pi)$.

\begin{proposition}[Budgeted Bellman Expectation]
\label{prop:bellman-expectation}
The value functions $V^\pi, Q^\pi$ of a budgeted policy $\pi$ verify:
\begin{align}
    V^\pi(\os) &= \sum_{\oa\in\ocA}\pi(\oa | \os) Q^\pi(\os, \oa),\\
    Q^\pi(\os, \oa) &= R(\os, \oa) + \gamma\sum_{\os'\in\ocS}\ov{P}\left(\os' \condbar \os, \oa\right) V^\pi(\os') \label{eq:bellman_expectation_Q}.
\end{align}
\end{proposition}

\begin{definition}[Budgeted Optimality]
We now come to the definition of budgeted optimality. We want an optimal budgeted policy to: (i)~respect the cost budget $\beta$, (ii)~maximise the rewards $R_r$, (iii)~in case of tie, minimise the costs $R_c$. To that end, we define the following sets for all $\os\in\ocS$:
\begin{itemize}
    \item Admissible policies: $\Pi_a(\os) \eqdef \{\pi\in\Pi: V_c^\pi(\os) \leq \beta\}$ where $\os = (s, \beta)$.
    \item Best admissible policies: $\Pi_r(\os) \eqdef \argmax_{\pi\in\Pi_a(\os)}  V_r^\pi(\os)$.
    \item Optimal policies: $\Pi^*(\os) \eqdef \argmin_{\pi\in\Pi_r(\os)}  V_c^\pi(\os)$.
\end{itemize}
\end{definition}

\begin{assumption}[Existence of optimal policy]
There exists one budgeted policy $\pi^*$ that uniformly maximises $V_r^\pi$ under constraint on $V_c^\pi$ for all augmented states $\os$.
\begin{align*}
    \exists \pi^*\in\Pi: \forall \os\in \ocS, \pi^*\in\Pi^*(\os)
\end{align*}
\end{assumption}

Let $\pi^*$ be one such optimal budgeted policy, we finally define the optimal budgeted value functions $V^*$, $Q^*$. For all $\os, \oa \in \ocS\times\ocA$:
\begin{equation*}
    V^*(\os) \eqdef V^{\pi^*}(\os), \qquad Q^*(\os, \oa) \eqdef  Q^{\pi^*}(\os, \oa),
\end{equation*}

\begin{remark}[Independence on the representative]
\label{rmk:independence}
Though they seemingly depend on the choice of $\pi^*$, the value functions $V^*$ and $Q^*$ are well-defined because all optimal policies in $\Pi^*$ share the same $V_r, Q_r$ as they belong to $\Pi_r$, and the same $V_c, Q_c$  as they belong to $\Pi^*$.
\end{remark}

\begin{theorem}[Budgeted Bellman Optimality]
\label{thm:bellman-optimality}
The optimal action-value of a budgeted policy verifies:
\begin{equation}
\label{eq:bellman-optimality}
    Q^{*}(\os, \oa) = \cT Q^{*}(\os, \oa) \eqdef R(\os, \oa) + \gamma \sum_{\os'\in\ocS} \ov{P}(\ov{s'} | \os, \oa)\sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q^*) Q^{*}(\ov{s'}, \ov{a'}),
\end{equation}
where the greedy policy $\pi_\text{greedy}$ is defined by: $\forall \os=(s,\beta)\in \ocS, \oa\in 
\ocA, \forall Q\in(\Real^2)^{\ocA\times\ocS},$
\begin{subequations}
\label{eq:pi_greedy}
\begin{equation}
    \pi_\text{greedy}(\oa|\os; Q) \in \argmin_{\rho\in\Pi_r^Q} \expectedvalueover{\oa\sim\rho}Q_c(\os, \oa), \label{eq:pi_greedy_cost}
\end{equation}
\begin{align}
    \text{where }\quad\Pi_r^Q \eqdef &\argmax_{\rho\in\cM(\ocA)} \expectedvalueover{\oa\sim\rho} Q_r(\os, \oa) \label{eq:pi_greedy_reward}\\
    & \text{ s.t. }  \expectedvalueover{\oa\sim\rho} Q_c(\os, \oa) \leq \beta. \label{eq:pi_greedy_constraint}
\end{align}
\end{subequations}
\end{theorem}

\begin{remark}[Appearance of the greedy policy]
\label{rmk:greedy}
In classical Reinforcement Learning, the greedy policy takes a simple form $\pi_\text{greedy}(s; Q^*) = \argmax_{a\in\cA} Q^*(s, a)$, and the term $\pi_\text{greedy}(a'|s';Q^*) Q^{*}(s', a')$ in \eqref{eq:bellman-optimality} conveniently simplifies to $\max_{a'\in \cA} Q^*(s', a')$. Unfortunately, in a budgeted setting the greedy policy requires solving the nested constrained optimisation program \eqref{eq:pi_greedy} at each state and budget in order to apply this Budgeted Bellman Optimality operator.
\end{remark}

\begin{remark}[Notations]
Because of the budget dynamics $\beta' = \beta_a$, we can see in \eqref{eq:bellman_expectation_Q} that for all $\os=(s, \beta)$ and $\oa=(a, \beta_a)$, $Q^\pi(\os, \oa)$ only depends on $s, a, \beta_a$ and not on $\beta$. We will slightly abuse notations and sometimes denote in the following $Q^\pi(s, a, \beta_a) \eqdef Q^\pi(\os, \oa)$.
\label{remark:notations}
\end{remark}

\paragraph{Budgeted Value Iteration}

The Budgeted Bellman Optimality equation is a fixed-point equation, which motivates the introduction of a fixed-point iteration procedure. We introduce Algorithm~\ref{algo:bvi}, a Dynamic Programming algorithm for solving known BMDPs.

\section{Budgeted Reinforcement Learning}

In this section, we consider BMDPs with unknown parameters that must be solved by interaction with an environment.

\subsection{Budgeted Fitted-Q}

When the BMDP is unknown, we need to adapt Algorithm~\ref{algo:bvi} to work with a batch of samples $\cD=\{(\os_i=(s_i,\beta_i),\oa_i=(a_i,\beta_{a_i}),R_i=({R_r}_i,{R_c}_i),\os_i'=(s_i',\beta_{a_i})\}_{i\in [0,N]}$ collected by interaction with the environment. Following \cref{remark:notations}, we rewrite this batch as  $\cD=\{s_i,a_i,\beta_{a_i},r_i,c_i,s_i'\}_{i\in [0,N]}$. Applying $\cT$ in \eqref{eq:bellman-optimality} would require computing an expectation $\expectedvalue_{\os'\sim \ov{P}}$ over next states $\os'$ and hence an access to the model $\ov{P}$. We instead use $\hat{\cT}$ in which this expectation is replaced by a sample average:
\begin{equation*}
    \hat{\cT} Q(\os, \oa, R, \os') \eqdef R + \gamma \sum_{\ov{a'}\in A} \pi_\text{greedy}(\ov{a'}|\ov{s'}; Q) Q(\ov{s'}, \ov{a'}).
\end{equation*}
We introduce in Algorithm~\ref{algo:bftq} the \emph{Budgeted-Fitted-Q} (BFTQ) algorithm, an extension of the \emph{Fitted-Q} (\FTQ) algorithm \citep{Ernst2005,Riedmiller2005} adapted to solve unknown BMDPs. Because we work with  continuous state space $\cS$ and budget space $\cB$, we need to employ function-approximation in order to generalise to nearby states and budgets. Precisely, given a parametrized model $Q_\theta$, we seek to minimise a regression loss $\cL(Q_\theta, Q_\text{target};\cD) = \sum_{\cD} ||Q_\theta(\os, \oa) - Q_\text{target}(\os, \oa, R, \os')||_2^2$.
Any model can be used, such as linear models, regression trees, or neural networks.

\begin{minipage}[t]{0.47\textwidth}
\vspace{0pt}  
\begin{algorithm}[H]
\label{algo:bvi}
\DontPrintSemicolon
\KwData{$P, R_r, R_c$}
\KwResult{$Q^*$}
$Q_{0} \leftarrow 0$\;
\Repeat{convergence}{
    $Q_{k+1} \leftarrow \cT Q_k$\;
}
\caption{Budgeted Value Iteration}
\end{algorithm}
\end{minipage}%
\hfill
\begin{minipage}[t]{0.47\textwidth}
\vspace{0pt}
\begin{algorithm}[H]
\label{algo:bftq}
\DontPrintSemicolon
\KwData{$\cD$}
\KwResult{$Q^*$}
$Q_{\theta_0} \leftarrow 0$\;
\Repeat{convergence}{
    $\theta_{k+1} \leftarrow \argmin_\theta \cL(Q_\theta, \hat{\cT} Q_{\theta_{k}}; \cD)$\;
}
\caption{Budgeted Fitted-Q}
\end{algorithm}
\end{minipage}

\subsection{Risk-sensitive exploration}

\label{sec:exploration}

In order to run Algorithm~\ref{algo:bftq}, we must first gather a batch of samples $\mathcal{D}$. Ideally we would need samples from the asymptotic state-budget distribution $\lim_{t\rightarrow\infty}\probability{\os_t}$ induced by an optimal policy $\pi^*$ given an initial distribution $\probability{\os_0}$, but as we are actually building this policy, it is not possible. Following the same idea of $\epsilon$-greedy exploration for \FTQ \citep{Ernst2005,Riedmiller2005}, we introduce an algorithm for risk-sensitive exploration. We follow an exploration policy: a mixture between a random budgeted policy $\pi_\text{rand}$ and the current greedy policy $\pi_\text{greedy}$. The batch $\cD$ is split into several mini-batches generated sequentially, and $\pi_\text{greedy}$ is updated by running Algorithm~\ref{algo:bftq} on $\cD$ upon mini-batch completion. $\pi_\text{rand}$ is designed to obtain trajectories that only explore feasible budgets: we impose that the joint distribution $\probability{a, \beta_a}$ verifies $\expectedvalue[\beta_a]\leq\beta$. This condition defines a probability simplex $\Delta_{\ocA}$ from which we sample uniformly. Finally, when interacting with an environment the initial state $s_0$ is usually sampled from a starting distribution $\probability{s_0}$. In the budgeted setting, we also need to sample the initial budget $\beta_0$. Importantly, we pick a uniform distribution $\probability{\beta_0} = \cU(\cB)$ so that the entire range of risk-level is explored, and not only reward-seeking behaviours as would be the case with a traditional risk-neutral $\epsilon$-greedy strategy. The pseudo-code of our exploration procedure is shown in Algorithm~\ref{algo:risk-sensitive-exploration} in the Supplementary Material.

\section{A Scalable Implementation}

\label{sec:scalable-bftq}
In this section, we introduce an implementation of the BFTQ algorithm designed to operate efficiently and handle large batches of experiences $\mathcal{D}$.

\subsection{How to compute the greedy policy?}

As stated in Remark~\ref{rmk:greedy}, computing the greedy policy $\pi_\text{greedy}$ in \eqref{eq:bellman-optimality} is not trivial since it requires solving the nested non-linear constrained optimisation program \eqref{eq:pi_greedy}. However, it can be solved efficiently by exploiting the \emph{structure} of the set of solutions with respect to $\beta$, which is concave and increasing.

\begin{proposition}[Equality of $\pi_\text{greedy}$ and $\pi_\text{hull}$]
\label{prop:bftq_pi_hull}
Algorithm~\ref{algo:bvi} and Algorithm~\ref{algo:bftq} can be run by replacing $\pi_\text{greedy}$ in the definition \eqref{eq:bellman-optimality} of $\cT$ with $\pi_\text{hull}$ as described in Algorithm~\ref{algo:pi_hull}.
\begin{equation*}
    \pi_\text{greedy}(\oa|\os; Q) = \pi_\text{hull}(\oa|\os; Q)
\end{equation*}
\end{proposition}

\begin{algorithm}
\DontPrintSemicolon
\KwData{$s$, $Q$}
$Q^+\leftarrow \{Q_c > \min \{Q_c(s,\oa) \text{ s.t. }\oa\in\argmax_{\oa} Q_r(s,\oa)\} \}$\tcp*[f]{dominated points}\;
$\cF \leftarrow \text{top frontier of }\texttt{convex\_hull}(Q(s,\ocA) \setminus Q^+)$\tcp*[f]{candidate mixtures}\;
$\cF_Q \leftarrow \cF\cap Q(s,\ocA)$\;
\For{points $q = Q(s,\oa)\in\cF_Q$ in clockwise order}{
\uIf{find two successive points $((q_c^1, q_r^1), (q_c^2, q_r^2))$ of $\cF_Q$ such that $q_c^1 \leq \beta < q_c^2$}{
$p \leftarrow (\beta - q_c^1) / (q_c^2 - q_c^1)$\;
\Return the mixture $(1-p)\delta(\oa-\oa^1) + p\delta(\oa-\oa^2)$\;
}}
\lElse{\Return $\delta(\oa - \argmax_{\oa} Q_r(s,\oa))$\tcp*[f]{Budget $\beta$ always respected}}
\caption{Convex hull policy $\pi_\text{hull}(\oa|\os; Q)$}
\label{algo:pi_hull}
\end{algorithm}

The computation of $\pi_\text{hull}$ in \cref{algo:pi_hull} is illustrated in \cref{fig:hull}.

\begin{minipage}[t]{0.5\textwidth}
\vspace{0pt}
\begin{figure}[H]
    \centering
    \includegraphics[width=\linewidth]{source/img/pi.pdf}
    \caption{Representation of $\pi_\text{hull}$. When the budget lies between $Q(\os,\oa_1)$ and $Q(\os,\oa_2)$, two points of the top frontier of the convex hull, then the policy is a mixture of these two points.}
    \label{fig:hull}
\end{figure}
\end{minipage}
\begin{minipage}[t]{0.5\textwidth}
\begin{figure}[H]
    \centering
    \input{source/architecture.tex}
    \caption{Neural Network for $Q$-functions approximation when $\cS=\Real^2$ and $|\cA| = 2$.}
    \label{fig:architecture}
\end{figure}
\end{minipage}

\subsection{Function approximation}

Neural networks are well suited to model Q-functions in Reinforcement Learning algorithms \citep{Riedmiller2005,Mnih2015}.  We approximate $Q = (Q_r, Q_c)$ using one single neural network. Thus, the two components are jointly optimised which accelerates convergence and fosters learning of useful shared representations. Moreover, as in \citep{Mnih2015} we are dealing with a finite (categorical) action space $\cA$, instead of including the action in the input we add the output of the $Q$-function for each action to the last layer. Again, it provides a faster convergence toward useful shared representations and it only requires one forward pass to evaluate all action values. Finally, beside the state $s$ there is one more input to a budgeted $Q$-function:~the budget $\beta_a$. This budget is a scalar value whereas the state $s$ is a vector of potentially large size. To avoid a weak influence of $\beta$ compared to $s$ in the prediction, we include an additional encoder for the budget, whose width and depth may depend on the application. A straightforward choice is a single layer with the same width as the state. The overall architecture is shown in Fig.~\ref{fig:architecture}.

\subsection{Parallel computing}

In a simulated environment, a first process that can be distributed is the collection of samples in the exploration procedure of Algorithm~\ref{algo:risk-sensitive-exploration}, as $\pi_\text{greedy}$ stays constant within each mini-batch which avoids the need of synchronisation between workers. Second, the main bottleneck of \BFTQ is the computation of the target $\cT Q$. Indeed, when computing $\pi_\text{hull}$ we must perform at each epoch a Graham-scan of complexity $O(|\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$ per sample in $\mathcal{D}$ to compute the convex hulls of $Q$. The resulting total time-complexity is $O(|\mathcal{D}|/(1-\gamma) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$. This operation can easily be distributed over several CPUs provided that we first evaluate the model $Q(s',\cA,\tilde{\cB})$ for each sample $s'\in\cD$, which can be done in a single forward pass. By using multiprocessing in the computations of $\pi_\text{hull}$, we enjoy a linear speedup.% and reduce the time-complexity to $O(|\mathcal{D}|/(W(1-\gamma)) |\cA\tilde{\cB}| \log |\cA\tilde{\cB}|)$.
The full description of our scalable implementation of \BFTQ is recalled in Algorithm~\ref{algo:bftq_full} in Supplementary Material~\ref{sec:bftq-full}.

\section{Experiments}

There are two hypotheses we want to validate.

\paragraph{Exploration strategies} We claimed in Section~\ref{sec:exploration} that a risk-sensitive exploration was required in the setting of BMDPs. We test this hypotheses by confronting our strategy to a classical risk-neutral strategy. The latter is chosen to be a $\epsilon$-greedy policy slowly transitioning from a random to a greedy policy\footnote{We train this greedy policy using \FTQ.} that aims to maximise $\expectedvalue_{\pi} G_r^\pi$ regardless of $\expectedvalue_{\pi} G_c^\pi$. The quality of the resulting batches $\cD$ is assessed by training a BFTQ policy and comparing the resulting performance.

\paragraph{Budgeted algorithms} We compare our  scalable \BFTQ algorithm described in Section~\ref{sec:scalable-bftq} to an \FTQl baseline. This baseline consists in approximating the BMDP by a finite set of CMDPs problems. We solve each of these CMDP using the standard technique of Lagrangian Relaxation: the cost constraint is converted to a soft penalty weighted by a Lagrangian multiplier $\lambda$ in a surrogate reward function: $\max_{\pi} \expectedvalue_{\pi} G_r^\pi - \lambda G_c^\pi\label{eq:lagrangian}$. The resulting MDP can be solved by any RL algorithm, and we chose $\FTQ$ for being closest to $\BFTQ$.
In these experiments, each run is composed of a training and a test phase and repeated $N_{\text{seeds}}$ times. Parameters of the algorithms can be found in Supplementary Material~\ref{sec:algorithms-parameters}

\subsection{Environments}

We evaluate our method on three different environments involving reward-cost trade-offs. Their parameters can be found in Supplementary Material~\ref{sec:env-parameters}

\paragraph{Corridors}
This simplistic environment is only meant to highlight clearly the specificity of exploration in a budgeted setting. It is a continuous gridworld with Gaussian perturbations, consisting in a maze composed of two corridors: a risky one with high rewards and costs, and a safe one with low rewards and no cost. In both corridors the outermost cell is the one yielding the most reward, which motivates a deep exploration.

\paragraph{Spoken dialogue system}
Our second application is a slot-filling simulation \citep{Li2009ReinforcementLF,chandramohan2010optimizing}. The system fills in a form of slot-values by interacting with a user through speech, before sending them a response. For example, in the restaurant domain, it may ask for three slots : the area of the restaurant, the price-range and the food type. The user could respectively provide those three slot-values : \texttt{Cambridge}, \texttt{Cheap} and \texttt{Indian-food}. In this application, we do not focus on how to extract such information from the user utterances, we rather focus on decision-making for filling in the form. To that end, the system can choose among a set of generic actions. As in \citep{carrara2018safe} there are two ways of asking for a slot value: a slot value can be either be provided with an utterance, which may to speech recognition errors with some probability, or by requiring the user to fill-in the slots by using a numeric pad. In this case, there are no recognition errors but a counterparty risk of hang-up: we assume that manually filling a key-value form is time-consuming and annoying. The environment yields a reward if all slots are filled without errors, and a constraint if the user hang-ups. Thus, there is a clear trade-off between using utterances and potentially committing a mistake, or using the numeric pad and risking a premature hang-up.
% Note that batch reinforcement learning has already been applied to spoken dialogue systems in \citep{Li2009ReinforcementLF,chandramohan2010optimizing} which makes this task a realistic candidate for optimisation with our method.

\paragraph{Autonomous driving}
In our third application, we use the \href{https://github.com/eleurent/highway-env}{highway-env} environment \citep{Leurent2018} for simulated highway driving and behavioural decision-making.
We propose a novel task that displays a clear trade-off between safety and efficiency. The agent controls a vehicle with a finite set of manoeuvres implemented by low-lever controllers: $\mathcal{A}$ = \{\text{no-op}, \text{right-lane}, \text{left-lane}, \text{faster}, \text{slower}\}. It is driving on a two-lane road populated with other traffic participants: the vehicles in front of the agent drive slowly, and there are incoming vehicles on the opposite lane. Their behaviours are randomised, which introduces some uncertainty with respect to their possible future trajectories.
The task consists in driving as fast as possible, which is modelled by a reward proportional to the velocity: $R_r(s_t, a_t) \propto v_t$. This motivates the agent to try and overtake its preceding vehicles by driving fast on the opposite lane. This optimal but overly aggressive behaviour can be tempered through a cost function that embodies a safety objective: $R_c(s_t, a_t)$ is set to $1/H$ whenever the ego-vehicle is driving on the opposite lane, where $H$ is the episode horizon. Thus, the constrained signal $G_c^\pi$ is the maximum proportion of time that the agent is allowed to drive on the wrong side of the road.

\subsection{Results}

In the following figures, each patch represents the mean and 95\% confidence interval over $N_{\text{seeds}}$ seeds of the means of $(G_r^\pi,G_c^\pi)$ over $N_\text{trajs}$ trajectories\footnote{Examples of policy execution can be found in Supplementary Material~\ref{sec:bftq-executions}}. That way, we display the variation related to learning (and batches) rather than the variation in the execution of the policies.

We first bring to light the role of risk-sensitive exploration in the \text{corridors} environment: Figure~\ref{fig:exploration} shows the set of trajectories collected by each exploration strategy, and the resulting performance of a budgeted policy trained on each batch. The trajectories (purple) in the risk-neutral batch (left) are concentrated along the risky corridor (orange) and ignore the safe corridor (green), which results in bad performances in the low-risk regime. Conversely, trajectories in the risk-sensitive batch (right) are well distributed among both corridors and the corresponding budgeted policy achieves good performance across the whole spectrum of risk budgets.
\begin{figure}[tp]
    \centering
    \includegraphics[width=0.23\textwidth]{source/img/risk-neutral.png}
    \includegraphics[width=0.23\textwidth]{source/img/risk-sensitive.png}
    \includegraphics[page=1, width=0.45\textwidth]{source/img/corridors}
    \caption{Trajectories (left) and performances (right) of two exploration strategies in the \text{corridors} environment. }
    \label{fig:exploration}
\end{figure}
% \begin{figure}[tp]
%     \begin{center}
%         \subfloat[\texttt{slot-filling}]{
%         \includegraphics[width=0.5\textwidth]{source/img/slot-filling}
%         \label{fig:slot-filling}
%         }
%         \subfloat[\texttt{highway-env}]{
%         \includegraphics[width=0.5\textwidth]{source/img/highway}
%         \label{fig:highway-env}
%         }
%         \caption{Performance comparison of \FTQl and BFTQ}
%         \label{fig:results}
%     \end{center}
% \end{figure}
\begin{figure}[tp]
    \begin{center}
    \includegraphics[width=0.49\linewidth]{source/img/slot-filling}
    \includegraphics[width=0.49\linewidth]{source/img/highway}
    \caption{Performance comparison of \FTQl and BFTQ on \text{slot-filling} (left) and \text{highway-env}(right) }
    \label{fig:results}
    \end{center}
\end{figure}

In a second experiment displayed in Fig.~\ref{fig:results}, we compare the performance of \FTQl to that of \BFTQ in the slot-filling and autonomous driving tasks. For each algorithm, we plot the reward-cost trade-off curve. In both cases, BFTQ performs almost as well as \FTQl despite only requiring a single model. All budgets are well-respected on \text{slot-filling}, but on \text{highway-env} we can notice in the bottom-left corner an underestimation of $Q_c$, since $\expectedvalue[G_c|\beta=0] \simeq 0.1 $. However, these estimates still provide a reasonable prior approximation of the expected cost of the policy, which cannot be obtained with \FTQl. In addition, BFTQ behaves more consistently than \FTQl overall, as shown by its lower extra-seed variance. Qualitatively, the budgeted agents display a wide variety of behaviours. In the \texttt{highway-env} experiment, when $\beta = 1$, the ego-vehicle drives in a very aggressive style: it immediately switches to the opposite lane and drives as fast as possible to pass slower vehicles, swiftly changing lanes to avoid incoming traffic. On the contrary when $\beta = 0$, the ego-vehicle is conservative: it stays on its lane and drives at a low velocity. With intermediate budgets such as $\beta = 0.2$, the agent sometimes decides to overtake its front vehicle but promptly steers back to its original lane afterwards.

\section{Discussion}

Algorithm~\ref{algo:bftq} is an algorithm for solving large unknown BMDPs with continuous states. To the best of our knowledge, there is no algorithm in the current literature that combines all those features.

%As a side note , utiliser Survey ~\citep{Roijers2013ASO} et  reformuler ça qui vient du papier  \citep{le2019batch} : Multi-objective Reinforcement Learning. Another related area is multi-objective reinforcement learning (MORL)(Van Moffaert & Nowe´, 2014; Roijers et al., 2013). Generally, research in MORL has largely focused on approximating the Pareto frontier that trades-off competing objectives (Van Moffaert & Nowe´, 2014; Roijers et al., 2013).The underlying approach to MORL frequently relies on linear or non-linear scalarization of rewards to heuristically turns the problem into a standard RL problem. Our proposed approach represents another systematic paradigm to solve MORL, whether in batch or online settings.}
%The literature of Multi-objective Reinforcement Learning mainly focuses on recovering the Pareto-optimal frontier,

Algorithms have been proposed for CMDPs, which are less flexible sub-problems of the more general BMDP. When the environment parameters ($P$, $R_r$, $R_c$) are known but not tractable, solutions relying on function approximation~\citep{Undurti} or approximate linear programming~\citep{Poupart2015} have been proposed. For unknown environments, batch algorithms exist in the tabular setting \citep{Thomas2015, Petrik2016, Laroche2017}. Online algorithms \citep{Geibel2005, Abe2010,ChowGJP15,AchiamHTA17} and a batch algorithm \citep{le2019batch} combine the best of both world by solving intractable unknown CMDPs.

A first way of solving a BMDP is to approximate it with a finite set of CMDPs. The solutions of these CMDPs take the form of mixtures between two deterministic policies\citep[Theorem 4.4][]{BEUTLER1985236}. To obtain these policies, one needs to evaluate their expected cost by interacting with environment\footnote{More details are provided in the Supplementary Material.}. Our solution not only requires one single model but also avoids any supplementary interaction.

The closest work to ours is the Dynamic Programming algorithm proposed by \citet{Boutilier_Lu:uai16} for solving BDMPs. However, their work was established for finite state spaces only, and their solution relies heavily on this property. For instance, they enumerate and sort the next states $s'\in\cS$ by their expected value-by-cost, which could not be performed in a continuous state space $\cS$. Moreover, they rely on the knowledge of the model ($P$, $R_r$, $R_c$), and do not address the question of learning from interaction data.

\section{Conclusion}
The BMDP framework is a principled framework for safe decision making under uncertainty, which could be beneficial to the diffusion of Reinforcement Learning in industrial applications. However, BMDPs could so far only be solved in finite state spaces which limits their interest in many use-cases. We extend their definition to continuous states by introducing of a novel Dynamic Programming operator, that we build upon to propose a Reinforcement Learning algorithm. In order to scale to large problems, we provide an efficient implementation that exploits the structure of the value function and leverages tools from Deep Distributed Reinforcement Learning. We show that on two practical tasks, our solution performs similarly to a baseline Lagrangian relaxation method while only requiring a single model to train.

%    \section*{Acknowledgments}
%    %
%    This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
%
%    Thanks to Guillaume Gautier, Fabrice Clerot, Xuedong Shang
\clearpage
\bibliographystyle{named}
\bibliography{budgeted_rl}

\clearpage
\input{supplementary.tex}

\end{document}