### 2019-06-11 poster/poster.tex
Add structure for poster

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% baposter Landscape Poster
% LaTeX Template
% Version 1.0 (15/5/13)
%
% Created by:
% Brian Amberg (baposter@brian-amberg.de)
%
% This template has been downloaded from:
% http://www.LaTeXTemplates.com
%
% License:
% CC BY-NC-SA 3.0 (http://creativecommons.org/licenses/by-nc-sa/3.0/)
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%----------------------------------------------------------------------------------------
%	PACKAGES AND OTHER DOCUMENT CONFIGURATIONS
%----------------------------------------------------------------------------------------

\documentclass[paperwidth=36in, paperheight=48in,portrait,fontscale=0.35, margin=2cm]{baposter}

\usepackage[utf8]{inputenc}
\usepackage[font=normalsize,labelfont=bf]{caption} % Required for specifying captions to tables and figures
\captionsetup{justification=centering}
\usepackage{booktabs} % Horizontal rules in tables
\usepackage{lmodern}
\usepackage{relsize} % Used for making text smaller in some places

% Other packages
\usepackage{tabularx}
\def\tabularxcolumn#1{m{#1}}
\def\imagetop#1{\vtop{\null\hbox{#1}}}
\usepackage{adjustbox}
\usepackage{subfig}
\usepackage{floatrow} % sidecapfloat
\usepackage{setspace}
\usepackage{xspace}
\usepackage{noReferences}
\usepackage{caption}
\usepackage{tikz}
\usepackage[customcolors]{hf-tikz}
%\usepackage{subcaption}
\usepackage{epsf}
\usepackage{amsmath}
\usepackage{mathtools}
\usepackage{amssymb}
\usepackage{amsfonts}
\usepackage{amsthm}
\usepackage{dsfont}
\usepackage{enumitem}
\usepackage{float}
\usepackage{cases}
\usepackage{verbatim}
\usepackage{array}
\usepackage{graphicx}
\usepackage{graphics}
\usepackage{multirow}
\usepackage{listings}
%\usepackage[authoryear]{natbib}
\usepackage{authblk}
\usepackage[linesnumbered,ruled,vlined]{algorithm2e}
\usepackage{eqparbox}
\usepackage{xcolor}
\usepackage[backend=biber, citestyle=authoryear, maxcitenames=2, maxbibnames=2]{biblatex}
\usepackage{lipsum}
\usepackage{svg}

\bibliography{../references.bib}

\input{style}
\input{../mathdef}

\begin{document}

%\background{ % Set the background to an image (background.pdf)
%\begin{tikzpicture}[remember picture,overlay]
%\draw (current page.north west)+(-2em,2em) node[anchor=north west]
%{\includegraphics[height=1.1\textheight]{background.png}};
%\end{tikzpicture}
%}
\begin{poster}{
grid=false,
borderColor=bordercol, % Border color of content boxes
headerColorOne=headercol1, % Background color for the header in the content boxes (left side)
headerColorTwo=headercol2, % Background color for the header in the content boxes (right side)
headerFontColor=headerfontcol, % Text color for the header text in the content boxes
boxColorOne=boxcolor, % Background color for the content in the content boxes
headershape=roundedright, % Specify the rounded corner in the content box headers
%headerfont=\Large\sf\bf, % Font modifiers for the text in the content box headers
headerfont=\Large\bf\textsc, %Sans Serif
textborder=rectangle,
background=none,
headerborder=open, % Change to closed for a line under the content box headers
boxshade=plain,
textfont={\setlength{\parindent}{0.0em}\sffamily},
headerheight={0.05\textheight},
eyecatcher=true
%columns=5
}
%
%----------------------------------------------------------------------------------------
%	Title and authors
%----------------------------------------------------------------------------------------
%
{
\includegraphics[width=14em]{./img/renault_group}
}
{
Practical Open-Loop Optimistic Planning
}
{
Edouard Leurent, Odalric-Ambrym Maillard
\vspace{-3\baselineskip}
}
{
\includegraphics[width=14em]{./img/inria_sc}
}

\setlength{\colheight}{0.92\textheight}

%----------------------------------------------------------------------------------------
%	Motivation
%----------------------------------------------------------------------------------------

\headerbox{\textsc{Motivation}}{name=motivation,span=1,column=0,row=0}{

}

%----------------------------------------------------------------------------------------
%	Setting
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Open-Loop Optimistic Planning}}{name=setting,span=1,column=0,row=0,below=motivation}{

}

%----------------------------------------------------------------------------------------
% Tools
%----------------------------------------------------------------------------------------
\headerbox{\textsc{What's wrong with OLOP?}}
{name=tools,column=0,below=setting,span =1}
{

}

%----------------------------------------------------------------------------------------
%	KL-OLOP
%----------------------------------------------------------------------------------------
\headerbox{\textsc{Kullback-Leibler OLOP}}{name=klolop,column=1,span=2}{

}

%----------------------------------------------------------------------------------------
%	Sample Complexity
%----------------------------------------------------------------------------------------
\headerbox{Sample complexity}
{name=sample,span=2,column=1,below=klolop}{ 
}

%----------------------------------------------------------------------------------------
%	Time and Memory Complexity
%----------------------------------------------------------------------------------------
\headerbox{Time and memory complexity}{name=time,span=2,column=1,below=sample}{

}

%
%%----------------------------------------------------------------------------------------
%% Experiments
%%----------------------------------------------------------------------------------------
%
\headerbox{Experiments}{name=experiments,span=2,column=1,below=time, above=bottom}{

}

%----------------------------------------------------------------------------------------
%	Acknowledgements
%----------------------------------------------------------------------------------------
% \headerbox{Acknowledgements}{name=ack,column=0,span=1,below=tools}{
% This work has been supported by CPER Nord-Pas de Calais/FEDER DATA Advanced data science and technologies 2015-2020, the French Ministry of Higher Education and Research, INRIA, and the French Agence Nationale de la Recherche (ANR).
% }

%----------------------------------------------------------------------------------------
%	References
%----------------------------------------------------------------------------------------
\headerbox{References}{name=refs,column=0,span=1,below=tools, above=bottom}{
    {
    \AtNextBibliography{\footnotesize}
    \setlength{\bibitemsep}{3pt}
    \printbibliography[heading=none]
    }
}

\end{poster}

\end{document}